---
title: "My Presentation"
author: "Shelby Bachman"
date: "7/15/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## My new slide

plain text

*italicized text*

**bold text**

`for (i %in% 1:10)`

*** 
A slide without a title

## My title

### A top level subheading

#### A second level subheading
<br> 
- Bullet 1
- Bullet 2

1. Item a
2. Item b

## A slide with code

Code + output:

```{r, comment = "", echo = TRUE}
head(mtcars)
```

Code only:

```{r, comment = "", eval = FALSE, echo = TRUE}
head(mtcars, 5)
```

## A slide with code (cont.)

Output only:

```{r, comment = "", echo = FALSE}
names(mtcars)
```

## A plot

```{r, echo = FALSE, fig.align = 'center', fig.cap = 'my figure'}
plot(mtcars$wt, mtcars$mpg, xlab = "weight", ylab = "MPG")
```

## Publishing RStudio Presentations

**Options:** 

- view in browser
- publish to RPubs (*use button in RStudio*)
- publish to Github (*need a gh-pages branch*)
  - then use this URL:`user.github.io/reponame/file.html`