---
title: "Week 3 Assignment"
subtitle: "Developing Data Products Course"
author: "Shelby Bachman"
date: "7/24/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(data.table)
library(dplyr)
library(tidyr)
library(plotly)

```

## Overview

In these slides, I will load, manipulate, and plot data to show the number of visitors to each of California's National Parks for each year from 2009 to 2018.

Data for this assignment was obtained from from the US National Park Service [Visitor Use Statistics](https://irma.nps.gov/Stats/) service.

## Load the data

First, I load the visitor data using the `fread()` function from the `data.table` package.

```{r load_data, echo = TRUE}

data <- fread('park_data.csv', header = TRUE)
head(data, 5)

```

## Tidy the data

 I would like to convert the data from wide to long format to thereby display the data with two variables, `year` and `visitors`. I do this below using the `gather()` function from the `tidyr` package.

```{r tidy_data, echo = TRUE}

data <- data %>%
  gather(key = year, value = visitors, -park) %>%
  mutate(park = as.factor(park),
         year = as.factor(year))
head(data, 3)

```

## Plot visitors by year using plotly

Next, I will plot the number of visitors (y) by year (x) as a line plot, with separate lines for each park. 

I will use the `plot_ly()` function within the `plotly` package for added interactivity. Find the final plot on the next page.

```{r plot_data, echo = TRUE, eval = FALSE}

plot_ly(data = data, x = ~year, y = ~visitors, type = "scatter", 
        mode = "lines", color = ~park)

```

## CA national park visitors by year

```{r plot_data_show, echo = FALSE, eval = TRUE}

plot_ly(data = data, x = ~year, y = ~visitors, type = "scatter", mode = "lines",
        color = ~park)

```

## Plot a subset of the data using ggplotly

Next, I will demonstrate the use of the `ggplotly()` function which converts a ggplot object to a plotly object.

I include code below that creates a ggplot object showing the number of visitors in each year from 2009-2018 for three of the lesser-visited parks, Channel Islands, Pinnacles, and Redwoods.

```{r plotly_data, echo = TRUE, eval = FALSE}

parksToKeep <- c('Channel Islands', 'Pinnacles', 'Redwoods')
p <- ggplot(data = data %>% 
              filter(park %in% parksToKeep),
            aes(x = year, y = visitors, colour = park)) +
  geom_point(size = 2) + geom_path(aes(group = park), size = 1) +
  scale_color_brewer(palette = "Accent") + 
  theme_minimal()
  
```

***

I then convert the ggpplot object to a plotly object:

```{r plotly_data_show, echo = FALSE, eval = TRUE, tidy=TRUE}

parksToKeep <- c('Channel Islands', 'Pinnacles', 'Redwoods')
p <- ggplot(data = data %>% 
              filter(park %in% parksToKeep),
            aes(x = year, y = visitors, colour = park)) +
  geom_point(size = 2) + geom_path(aes(group = park), size = 1) +
  scale_color_brewer(palette = "Accent") + 
  theme_minimal()

```

```{r plotly_data_show2, echo = TRUE, eval = TRUE}

ggplotly(p)

```
